
## Job satisfaction

Guardando alla tua domanda di OSP e cose che potrebbero essere divertenti, la variabile  latente per eccellenza e' la Job satisfaction. Tra quella per i soldi e per sicurezza (jbmssec) mi sembra ci sia piu' variabilita' per Job satisfaction about pay (jbmspay) e potremmo vedere:
 
1. dall'inizio HILDA, quali sono i lavori dove chi lavora e' piu' felice? [jbmo62]
2. ci sono settori che sono sistematicamente over-paid (qui dovresti vedere in primis mining e finanza. Anche il settore pubblico non e' male)
3. sono piu' felici uomi e donne? [sex]
4. immigrati o nativi? [ancob3]
5. giovani o vecchi? (possibile scala: 25-29; 30-39; 40-49; 50-59; 60+) [hgage + I(hgage^2)]
6. laureati o no? [edcoq --> ...]
7. lavoratori fedeli al'employer o switchers? (tenure: jbocct) 
8. in che stato? (hhstate)
9. e in che tipo di famiglia? (hhtype)
 
detto questo, mi sembra che il funzionale che tu stai usando e' legato al tempo: puo' essere usato anche su intensita', tipo prestigio (jbmoccs): il prestigio ha molte categorie ma e' stato surveyed solo in wave 5. Il che vuol dire mappare le occupation/Prestige scale da wave 5 e 'spararla' alle occupations di tutte le rimanenti waves.

#### Satisfaction for job pay

```{r subset, eval=F, include=T, echo=F, comment=NA, cache=F}
xsub=x[,c("xwaveid","hgsex","jbmspay","jbmssec","jbmo62","ancob3","hgage","ehtse","edcoq","jbocct","hhstate","hhtype","wsfei","wages_perc", "ehtse_au","ehtse_anglo","ehtse_others")]
xsub <- xsub[substr(as.character(xsub$jbmspay),1,1)!="-" & substr(as.character(xsub$jbmssec),1,1)!="-",]
xsub <- xsub[substr(xsub$hhstate,1,2)!="[-",]
xsub=droplevels(xsub)
```

Comparison of the effects on satisfation by salary vs wages_perc.

##### Salary:
```{r satisfaction_job_pay_salary, eval=F, include=T, echo=F, comment=NA, cache=F}
fit <- clm(jbmspay ~ hgsex + hgage + ancob3+ hhtype+ jbocct + I(log(wsfei)), data=xsub)
summary(fit)
extractAIC(fit)
```

##### wages_perc

```{r satisfaction_job_pay_perc, eval=F, include=T, echo=F, comment=NA, cache=F}
fitp <- clm(jbmspay ~ hgsex + hgage + ancob3+ hhtype+ jbocct + wages_perc, data=xsub)
summary(fitp)
extractAIC(fitp)
#fit <- clm(jbmspay ~ hgsex + jbmo62 + ancob3 + hgage + edcoq + jbocct + hhstate + hhtype + wages_perc, data=xsub)
#summary(fit)
#fitm <- clmm(jbmspay ~ hgsex + jbmo62 + ancob3 + hgage + edcoq + jbocct + hhstate + hhtype + wages_perc + (1|xwaveid), data=xsub)
```

The model fitted with the wager_perc variable is a better choice than that with the annula salary wsfei.


```{r satisfaction_job_pay_perc_interaction, eval=F, include=T, echo=F, comment=NA, cache=F}
fitp2 <- clm(jbmspay ~ hgsex + hgage +  hhtype+ ancob3 + jbocct:ancob3 + wages_perc:ancob3, data=xsub)
summary(fitp2)
extractAIC(fitp2)
```

AIC improved.

```{r satisfaction_job_pay_perc_interaction_rnd, eval=F, include=T, echo=F, comment=NA, cache=F}
fitp2rnd <- clmm(jbmspay ~ hgsex + hgage +  hhtype+ ancob3 + jbocct:ancob3 + wages_perc:ancob3 + (1|xwaveid), data=xsub)
summary(fitp2rnd)
extractAIC(fitp2rnd)
```

##### Smoothing satisfaction pay

```{r libraries-detach-attach, echo=F, cache=F}
detach("package:mgcv")
library(VGAM)
```

We now create three variables: the wage (or wage_perc) for Australians, English-speaking migrants and other migrants.
We fit a generalized additive model with the effect of these three variable smoothed.

```{r vgam1, eval=F, include=T, echo=F, comment=NA}
#fitgam0 <- vgam(jbmspay ~ hgsex + hgage + ancob3+ hhtype+ jbocct + s(wages_perc), data=xsub, family=propodds)
#plotvgam(fitgam0, rugplot=F)

##wage
xsub$w_au = xsub$w_anglo = xsub$w_others = 0
xsub$w_au[xsub$ancob3=="australians"] = xsub$wsfei[xsub$ancob3=="australians"]
xsub$w_anglo[xsub$ancob3=="imm_anglo"] = xsub$wsfei[xsub$ancob3=="imm_anglo"]
xsub$w_others[xsub$ancob3=="imm_others"] = xsub$wsfei[xsub$ancob3=="imm_others"]
fitgam_wage <- vgam(jbmspay ~ hgsex + hgage + hhtype+ jbocct + s(w_au) + s(w_anglo) + s(w_others), data=xsub, family=propodds)
plotvgam(fitgam_wage, se=T, rugplot=F, xlab="wage percentiles", ylab="effect on satisfaction", which.term=c("s(w_au)"),control=plotvgam.control(ylim=c(-.2,.4), lcol="blue", scol="blue"))
plotvgam(fitgam_wage, se=T, rugplot=F, which.term=c("s(w_anglo)"),control=plotvgam.control(add.arg=T, lcol="red", scol="red"))
plotvgam(fitgam_wage, se=T, rugplot=F, which.term=c("s(w_others)"),control=plotvgam.control(add.arg=T, lcol="black", scol="black"))
legend("bottomright",c("Australians", "English-speaking migrants", "Other migrants") , col=c("blue","red","black"), lty=c(1,1,1) )
```

The confidence intervals are too big, the wage is not the right variable to use.
We can now use the percentiles of wage in each job code.

```{r vgam2, eval=F, include=T, echo=F, comment=NA}
##wage_perc
xsub$wp_au = xsub$wp_anglo = xsub$wp_others = 0
xsub$wp_au[xsub$ancob3=="australians"] = xsub$wages_perc[xsub$ancob3=="australians"]
xsub$wp_anglo[xsub$ancob3=="imm_anglo"] = xsub$wages_perc[xsub$ancob3=="imm_anglo"]
xsub$wp_others[xsub$ancob3=="imm_others"] = xsub$wages_perc[xsub$ancob3=="imm_others"]
fitgam <- vgam(jbmspay ~ hgsex + hgage + hhtype+ jbocct + s(wp_au) + s(wp_anglo) + s(wp_others), data=xsub, family=propodds)
plotvgam(fitgam, rugplot=F, se=T, xlab="wage percentiles", ylab="effect on satisfaction", which.term=c("s(wp_au)"),control=plotvgam.control(ylim=c(-.3,.5), lcol="blue", scol="blue"))
plotvgam(fitgam, rugplot=F, se=T, which.term=c("s(wp_anglo)"),control=plotvgam.control(add.arg=T, lcol="red", scol="red"))
plotvgam(fitgam, rugplot=F, se=T, which.term=c("s(wp_others)"),control=plotvgam.control(add.arg=T, lcol="black", scol="black"))
legend("bottomright",c("Australians", "English-speaking migrants", "Other migrants") , col=c("blue","red","black"), lty=c(1,1,1) )
```

The results are interesting: [interpretation here].
Let's now study the satisfaction for pay as function of yearssince studying:

```{r vgam3, eval=F, include=T, echo=F, comment=NA}
fitgam <- vgam(jbmspay ~ hgsex + hgage + hhtype+ jbocct + ancob3 + s(wages_perc) + s(ehtse_au) + s(ehtse_anglo) + s(ehtse_others), data=xsub, family=propodds)
#Wages_perc
plotvgam(fitgam, rugplot=F, se=T, xlab="wage percentiles", ylab="effect on satisfaction", which.term=c("s(wages_perc)"),control=plotvgam.control(ylim=c(-.3,1), lcol="blue", scol="black"))
#Years since studying
plotvgam(fitgam, rugplot=F, se=F, xlab="Years since studying", ylab="effect on satisfaction", which.term=c("s(ehtse_au)"),control=plotvgam.control(ylim=c(-1,1), lcol="blue", scol="blue"))
plotvgam(fitgam, rugplot=F, se=F, which.term=c("s(ehtse_anglo)"),control=plotvgam.control(add.arg=T, lcol="red", scol="red"))
plotvgam(fitgam, rugplot=F, se=F, which.term=c("s(ehtse_others)"),control=plotvgam.control(add.arg=T, lcol="black", scol="black"))
legend("bottomright",c("Australians", "English-speaking migrants", "Other migrants") , col=c("blue","red","black"), lty=c(1,1,1) )
```

There are no differences between Australians, English-speaking migrant and other migrants in the satisfaction for pay as a function of the time since last studying.




#### Satisfaction for job security:

##### Salary:

```{r satisfaction_job_security_salary, eval=F, include=T, echo=F, comment=NA, cache=F}
fit <- clm(jbmssec ~ hgsex + hgage + ancob3+ hhtype+ jbocct + I(log(wsfei)), data=xsub)
summary(fit)
#fit <- clm(jbmssec ~ hgsex + hgage + ancob3+ hhtype+ jbocct + wages_perc + hhstate, data=xsub)
#summary(fit)
#Model is nearly unidentifiable
#fit <- clm(jbmssec ~ hgsex + hgage + ancob3+ hhtype+ jbocct+ edcoq + hhstate, data=xsub)
```

##### wages_perc

```{r satisfaction_job_security_perc, eval=F, include=T, echo=F, comment=NA, cache=F}
fit <- clm(jbmssec ~ hgsex + hgage + ancob3+ hhtype+ jbocct + wages_perc, data=xsub)
summary(fit)
```

No problems with this fit. wages_perc works better than wsfei to explain job satisfaction.


##### Smoothing satisfaction job security

We now create three variables: the wage (or wage_perc) for Australians, English-speaking migrants and other migrants.
We fit a generalized additive model with the effect of these three variable smoothed.

```{r vgam4, eval=F, include=T, echo=F, comment=NA}
library(VGAM)
#fitgam0 <- vgam(jbmssec ~ hgsex + hgage + ancob3+ hhtype+ jbocct + s(wages_perc), data=xsub, family=propodds)
#plotvgam(fitgam0, rugplot=F)

##wage
fitgam_wage <- vgam(jbmssec ~ hgsex + hgage + hhtype+ jbocct + s(w_au) + s(w_anglo) + s(w_others), data=xsub, family=propodds)
plotvgam(fitgam_wage, se=T, rugplot=F, xlab="wage percentiles", ylab="effect on satisfaction", which.term=c("s(w_au)"),control=plotvgam.control(ylim=c(-.2,.4), lcol="blue", scol="blue"))
plotvgam(fitgam_wage, se=T, rugplot=F, which.term=c("s(w_anglo)"),control=plotvgam.control(add.arg=T, lcol="red", scol="red"))
plotvgam(fitgam_wage, se=T, rugplot=F, which.term=c("s(w_others)"),control=plotvgam.control(add.arg=T, lcol="black", scol="black"))
legend("bottomright",c("Australians", "English-speaking migrants", "Other migrants") , col=c("blue","red","black"), lty=c(1,1,1) )
```

The confidence intervals are too big, the wage is not the right variable to use.
We can now use the percentiles of wage in each job code.

```{r vgam5, eval=F, include=T, echo=F, comment=NA}
##wage_perc
fitgam <- vgam(jbmssec ~ hgsex + hgage + hhtype+ jbocct + s(wp_au) + s(wp_anglo) + s(wp_others), data=xsub, family=propodds)
dev.off()
plotvgam(fitgam, rugplot=F, se=T, xlab="wage percentiles", ylab="effect on satisfaction", which.term=c("s(wp_au)"),control=plotvgam.control(ylim=c(-.3,.5), lcol="blue", scol="blue"))
plotvgam(fitgam, rugplot=F, se=T, which.term=c("s(wp_anglo)"),control=plotvgam.control(add.arg=T, lcol="red", scol="red"))
plotvgam(fitgam, rugplot=F, se=T, which.term=c("s(wp_others)"),control=plotvgam.control(add.arg=T, lcol="black", scol="black"))
legend("bottomright",c("Australians", "English-speaking migrants", "Other migrants") , col=c("blue","red","black"), lty=c(1,1,1) )
```

The results are interesting: [interpretation here].
Let's now study the satisfaction for pay as function of yearssince studying:

```{r vgam6, eval=F, include=T, echo=F, comment=NA}
fitgam <- vgam(jbmssec ~ hgsex + hgage + hhtype+ jbocct + s(wages_perc) + s(ehtse_au) + s(ehtse_anglo) + s(ehtse_others), data=xsub, family=propodds)
#Wages_perc
plotvgam(fitgam, rugplot=F, se=T, xlab="wage percentiles", ylab="effect on satisfaction", which.term=c("s(wages_perc)"),control=plotvgam.control(ylim=c(-.3,1), lcol="blue", scol="black"))
#Years since studying
plotvgam(fitgam, rugplot=F, se=F, xlab="Years since studying", ylab="effect on satisfaction", which.term=c("s(ehtse_au)"),control=plotvgam.control(ylim=c(-1,1), lcol="blue", scol="blue"))
plotvgam(fitgam, rugplot=F, se=F, which.term=c("s(ehtse_anglo)"),control=plotvgam.control(add.arg=T, lcol="red", scol="red"))
plotvgam(fitgam, rugplot=F, se=F, which.term=c("s(ehtse_others)"),control=plotvgam.control(add.arg=T, lcol="black", scol="black"))
legend("bottomright",c("Australians", "English-speaking migrants", "Other migrants") , col=c("blue","red","black"), lty=c(1,1,1) )
```



## Differences
TODO

```{r eval=F, include=T, echo=F, comment=NA}
d=deltapay=by(x, x$xwaveid, function(x){n=nrow(x);return(x$wsfei[2:n]/x$wsfei[1:(n-1)])})
plot(log(d[[1]]), xlim=c(0,12),ylim=c(0.1,20), xlab='wave', ylab='log(annual salary)')
for (i in 2:length(d)) {
  if (!is.null(d[[i]])){
    if (all(!is.na(d[[i]]))) {
      points(log(d[[i]]))
    }
  }
}

d=deltaperc=by(x, x$xwaveid, function(x){n=nrow(x);return(x$wages_perc[2:n]-x$wages_perc[1:(n-1)])})
plot(log(d[[1]]), xlim=c(0,12),ylim=c(0,1), xlab='wave', ylab='wages_perc')
for (i in 2:length(d)) {
  if (!is.null(d[[i]])){
    if (all(!is.na(d[[i]]))) {
      points(d[[i]])
    }
  }
}

```


































