---
output:
  html_document:
    keep_md: yes
---
```{r set-options, echo=F, cache=F}
options(width = 120)
```

```{r load-libraries, echo=F, cache=F}
library(mgcv)
library(ordinal)
```

# HILDA

Le variabili di Interesse sono edhigh1 (education) e jbmo62 (occupation). 
Come vedi ci sono relativamente pochi immigrati. 
Ti ho inviato un'extraction del pannello (xwaveid e wave). Ci sono 12 osservazioni al massimo ma come vedrai non per tutti.
I salari sono wscei (current weekly wage) e wsfei (Financial year wage) 

Covariates:
ancob: Country of birth
yrivwfst: year first interview
yrivwlst: year last interview
hh*: household...
edcqtyp: Currently studying full or part time
esdtl: Labour force status - detail
esempst: Current employment status
jb*: job... (jbmo62)
helth: Long term health condition
fmfsch: How much schooling father completed
fmmsch: How much schooling mother completed
anlote: Speak language other than English
aneab: How well speaks English
anyoa: Year first came to Australia to live (no anyoan)
anpappn: Australian visa - Primary applicant
anengfn: Is English the first language you learned to speak as a child
edcoqn: Country completed highest education
rtcompn: Retired completely from the workforce
rtyrn: Year retired
jbmoccs: occupational status scale, current main job
tcr: Number of own resident children
tcnr: Number of own non-resident children
hhyng: Age of youngest person in household
hhold: Age of oldest person in household
anmigc: Migration category when you or your family first arrived in Australia: **mostly missings**
fmnsib: How many siblings
edqenr: Ever enrolled in a course of study to obtain a qualification
edcoq: Country completed highest qualification
hgage: Age last birthday at June 30 2012
ehtse: Time since FT (full-time) education - years

## Data

Load the data and make some data managing: 

Cleaning:
1.  jbmo62: consolidate negative codes (different causes for missing info)
2.  ancob [country of birth] --> ancob3. 3 categories: AU, anglo (UK, Ireland, South Africa, Canada, USA), others
3.  edcoq3 [country completed highest education]: AU, anglo, others
4.  edqenr [Ever enrolled in a course of study to obtain a qualification]: consolidate neg. codes and missings 
5.  aneab [How well speaks English]: put "very well" for australians
6.  Date of birth <- 2000 + x$wave - x$hgage
7.  anyoa [year of arriva] <-- dob for australian born
8.  Factors; ordered factors: ("xwaveid","edcoq","edqenr","ancob3","aneab"; "jbmspay","jbmssec")
9.  annual salary wsfei=0 --> 1 (for log)
10. Household income contribution from others (hifefp-hifefn-wsfei)
11. Household financial year gross (hifefp)=0 --> 1
New variables:
12. years since studying FT stratified by country of birth
13. wave stratified by age groups, age_limits = c(15,25,35,45,55)
14. wave stratified by migration status (Australians vs English speaking migrants vs other migrants)
15. wave stratified by labour force status (esdtl)
16. wave straified by current employment status (esempst)
17. wave stratified by education
18. wave stratified by skilled jobs (3 categories)
19. wave stratified by subject median salary
Cleaning:
20. delete records of people working 997 hours per week (~3500 records)
New variable:
21. wages_perc: for each job in jbmo62, it's the percentile in the distribution of salaries across all the years
Filtering:
22. 18<= age <=65
23. Only record when questionnarie has been filled



```{r data_manipulation, eval=T, include=T, echo=F, comment=NA, cache=F}
#library(readstata13)
#setwd("~/Dropbox/Documents/Research/Migrations/data")
#load("Hilda.RData")
#x=read.dta13("MAURIZIO_HILDA_extraction.dta")
library(Migrations)
x = clean_data(HILDA)
```

## Models suggested by Max:

1. wage = costant + demographics + experience(age-anni scuola) + caract.employer(if available) + location + e
2. participation (0=no,1=yes) = costant + demographics + education + young children + wealth + e

### Variables

* wscei: weekly wage ---> replaced with wsfei (annual wage)
~
* ancob  [immigrato o no] --> replaced with ancob3: 3 categories: au, anglo, others
* edcoq  [education in au]
* edqenr [qualifica in au]
* fmfo61 and/or fmmo61 [figlio di operai]
...controlli...
* aneab [conoscenze inglese: missings born in au replaced with 1:very well]
* anyoa [year of arrival]. For non-migrants, replace with 2012-age_2012
* edagels o edhigh1 [years of education] ---> used edhigh1
* esdtl [participation to labour market]
* hgage [age]
* hgsex [sex]
* marital status [mrcurr]
* hhmsr o hhstate [location]
* jbmo62 [occupation]
* jbmwpsz [no employees]
* jbocct [tenure current job]
* ehtse [years since studying FT]

### Analyses 

In fit26, we fit the number of hours worked against several covs.
In the next analyses, we will fit identical models with different dependent variables: 
1. fit27: wsfei (annual salary); 
2. fit27weekly: wscei (weekly salary); 
3. fit27hrwage: hrw (hourly salary) 
4. fit28: wage_perc (the annual salary is transformed into the percentile specific to the subject's occupation).
5. In the third fit (fit29), we add the random effects over the subjects (the fit fails when using the annual salary as dependent variable, but succesfully converge with wages_perc).

```{r fit26, eval=T, include=T, echo=T, comment=NA, cache=F}
model26=(jbhru ~ s(wave) +  hgage + I(hgage^2) + sex + ancob3 + edhigh1 + jbmo62 + edcoq3 + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct  + other_household)
fit26 = gam(model26, data=x)
anova(fit26)
plot(fit26,pages=1, scale=-1, rug=F)
```

There is no indication of any specific effect of the global crisis on the number of hours worked per week.

```{r fit27, eval=T, include=T, echo=T, comment=NA, cache=F}
model27=(log(wsfei) ~ s(wave) +  hgage + I(hgage^2) + sex + ancob3 + edhigh1 + jbmo62 + edcoq3 + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct + aneab + other_household)
fit27 = gam(model27, data=x)
anova(fit27)
plot(fit27,pages=1, scale=-1, rug=F)
```



```{r fit27bis, eval=T, include=T, echo=T, comment=NA, cache=F}
model27bis=(log(wsfei) ~ s(wave) +  hgage + I(hgage^2) + sex + ancob3 + edhigh1 + jbmo62 + edcoq3 + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct + jbhru  + other_household)
fit27bis = gam(model27bis, data=x)
anova(fit27bis)
plot(fit27bis,pages=1, scale=-1, rug=F)
```

In fit27bis we repeat the same regression as in fit27, but consider the weekly worked hours as a covariate (significant). The effect of wave over the salary is similar as in fit27.

```{r fit27weekly, eval=T, include=T, echo=T, comment=NA, cache=F}
x2=x[x$wscei>0 & x$jbhru>0,]
x2$wpw = x2$wscei/x2$jbhru
model27weekly=(log(wscei) ~ s(wave) +  hgage + I(hgage^2) + sex + ancob3 + edhigh1 + jbmo62 + edcoq3 + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct  + other_household)
fit27weekly = gam(model27weekly, data=x2)
anova(fit27weekly)
plot(fit27weekly,pages=1, scale=-1, rug=F)
```

The global crisis effect is obvious also using the weekly salary.

```{r fit27hrwage, eval=T, include=T, echo=T, comment=NA, cache=F}
x2=x[x$wscei>0 & x$jbhru>0,]
x2$hrw = x2$wscei/x2$jbhru
model27hrwage=(hrw ~ s(wave) +  hgage + I(hgage^2) + sex + ancob3 + edhigh1 + jbmo62 + edcoq3 + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct  + other_household)
fit27hrwage = gam(model27hrwage, data=x2)
anova(fit27hrwage)
plot(fit27hrwage,pages=1, scale=-1, rug=F)
```

No effect of the global crisis on the hourly wage. This has nothing to do with the status of native or migrant (ancob3) or the knowledge of English (aneab)



```{r fit28, eval=T, include=T, echo=T, comment=NA, cache=F}
model28=(wages_perc ~ s(wave) +  hgage + I(hgage^2) + sex + ancob3 + edhigh1 + jbmo62 + edcoq3 + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct + other_household)
fit28 = gam(model28, data=x)
anova(fit28)
plot(fit28,pages=1, scale=-1, rug=F)
```


The last fit (fit28, rnd effects on xwaveid) shows that the effect on the wave can be assumed linear and a few covariates are no more significant and can be eliminated from the model. These are:

* Occupation of parents
* ancob3: australian, migrant anglo, migrant other.

This result can be interpreted considering that the dependent variable is the salary percentile of the subject specific to his job. The occupation of the parents and the status of migrant could contribute to choice of the job, but not to the salary in it (hypothesis to test).

# Variations on fit27

## fit27: random effects

```{r fit27rnd, eval=F, include=T, echo=T, comment=NA, cache=F}
#load("~/Dropbox/Attachments/-.Migrations/data/fit27rnd.RData")
model27rnd=(log(wsfei) ~ s(wave) +  hgage + I(hgage^2) + sex + ancob3 + edhigh1 + jbmo62 + edcoq + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct + aneab)
fit27rnd = gamm(model27, random=list(xwaveid=~1), data=x)
anova(fit27rnd$gam)
plot(fit27rnd$gam,pages=1, scale=-1, rug=F)
```


## fit27: wave stratified by age groups

```{r fit27byAge, eval=F, include=T, echo=T, comment=NA, cache=F}
model27byAge=(log(wsfei) ~ s(wave, age_category, bs="fs",k=5) +  hgage + I(hgage^2) + sex + ancob3 + edhigh1 + jbmo62 + edcoq3 + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct + other_household)
fit = gam(model27byAge, data=x)
snames=fit$smooth[[1]]$flev
anova(fit)
plot(fit,pages=1, scale=-1, rug=F) #all.terms=T
legend("bottomleft",snames,  lty=1:length(snames))
```

The effect of the global crisis has been similar in all the 5 age groups.


## fit27: wave stratified by migration status (Australians vs English speaking migrants vs other migrants)

```{r fit27byMigrationStatus, eval=F, include=T, echo=T, comment=NA, cache=F}
model27byMigrationStatus=(log(wsfei) ~ s(wave, ancob3, bs="fs",k=5) + hgage + I(hgage^2) + sex + ancob3 + edhigh1 + jbmo62 + edcoq3 + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct + other_household) # aneab cannot stay in the regression as few unique covariate combinations with wave_*
fit = gam(model27byMigrationStatus, data=x)
snames=fit$smooth[[1]]$flev
anova(fit)
plot(fit,pages=1, scale=-1, rug=F) #all.terms=T
legend("bottomleft",snames,  lty=1:length(snames))
```

No differences in migration status.

## fit27: wave stratified by labour force status 

```{r fit27byLabourForceStatus, eval=F, include=T, echo=T, comment=NA, cache=F}
model27byLabourForceStatus=(log(wsfei) ~ s(wave, esdtl, bs="fs",k=5)+ hgage + I(hgage^2) + sex + ancob3 + edhigh1 + jbmo62 + edcoq3 + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct + other_household) 
fit = gam(model27byLabourForceStatus, data=x)
snames=fit$smooth[[1]]$flev
anova(fit)
plot(fit,pages=1, scale=-1, rug=F) #all.terms=T
legend("bottomleft",snames,  lty=1:length(snames))
```

## fit27: wave stratified by employment status 

```{r fit27byEmploymentStatus, eval=F, include=T, echo=T, comment=NA, cache=F}
model27byEmploymentStatus=(log(wsfei) ~ s(wave, esempst, bs="fs",k=5) + hgage + I(hgage^2) + sex + ancob3 + edhigh1 + jbmo62 + edcoq3 + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct + other_household) 
fit = gam(model27byEmploymentStatus, data=x)
snames=fit$smooth[[1]]$flev
anova(fit)
plot(fit,pages=1, scale=-1, rug=F) #all.terms=T
legend("bottomleft",snames,  lty=1:length(snames))
```


## fit27: wave stratified by education

```{r fit27byEducation, eval=F, include=T, echo=T, comment=NA, cache=F}
model27byEducation=(log(wsfei) ~ s(wave, edhigh1, bs="fs",k=5) + hgage + I(hgage^2) + sex + ancob3 + edhigh1 + jbmo62 + edcoq3 + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct + other_household)
fit = gam(model27byEducation, data=x)
snames=fit$smooth[[1]]$flev
anova(fit)
plot(fit,pages=1, scale=-1, rug=F) #all.terms=T
legend("bottomleft",snames,  lty=1:length(snames))
```

There is a weak effect: the higher the education level, the lighter the effect of the global crisis.

## fit27: wave stratified by skilled jobs (3 categories)

```{r fit27bySkilledJobs, eval=F, include=T, echo=T, comment=NA, cache=F}
model27bySkilledJobs=(log(wsfei) ~ s(wave, skill_category, bs="fs",k=5) + hgage + I(hgage^2) + sex + ancob3 + edhigh1 + jbmo62 + edcoq3 + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct + other_household)
fit = gam(model27bySkilledJobs, data=x)
snames=fit$smooth[[1]]$flev
anova(fit)
plot(fit,pages=1, scale=-1, rug=F) #all.terms=T
legend("bottomleft",snames,  lty=1:length(snames))
```

No differences in job skill categories.


## fit27: wave stratified by subject median salary (5 categories: 0<5000<50000<100000<200000+)

```{r fit27bySalary, eval=F, include=T, echo=T, comment=NA, cache=F}
model27bySalary=(log(wsfei) ~ s(x$wave, wage_category, bs="fs",k=5) + hgage + I(hgage^2) + sex + ancob3 + edhigh1 + jbmo62 + edcoq3 + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct + other_household)
fit = gam(model27bySalary, data=x)
snames=fit$smooth[[1]]$flev
anova(fit)
plot(fit,pages=1, scale=-1, rug=F) #all.terms=T
legend("bottomleft",snames,  lty=1:length(snames))
```



# Variations on fit28

## fit28: random effects

```{r fit28rnd, eval=F, include=T, echo=F, comment=NA}
#load("~/Dropbox/Attachments/-.Migrations/data/fit28rnd.RData")
fit28rnd = gamm(model28, random=list(xwaveid=~1), data=x)
#summary(fit28rnd$lme)
anova(fit28rnd$gam)
plot(fit28rnd$gam,pages=1)

#Males vs females
#xm=x[x$sex=="[1] Male",]
#xf=x[x$sex=="[2] Female",]
#model26m=(wsfei ~ s(wave) +  hgage + I(hgage^2) + s(anyoa) + ancob3 + edhigh1 + jbmo62 + edcoq + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct + aneab)
#fit26m = gam(model26m, data=xm)
#anova(fit26m)
#plot(fit26m,pages=1)
#model26f=(wsfei ~ s(wave) +  hgage + I(hgage^2) + s(anyoa) + ancob3 + edhigh1 + jbmo62 + edcoq + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + #jbmwpsz + jbocct + aneab)
#fit26f = gam(model26f, data=xf)
#anova(fit26f)
#plot(fit26f,pages=1)
```



# Family

```{r family_wage, eval=T, include=T, echo=F, comment=NA}
model30=(log(hifefp) ~ s(wave) + ancob3 + edhigh1 + jbmo62 + edcoq3 + edqenr +  esdtl + hhstate + jbmwpsz + jbocct + other_household)
fit30 = gam(model30, data=x)
anova(fit30)
plot(fit30,pages=1, scale=-1, rug=F)

```

## More on time-related covariates


Let's consider the years since studying FT (ehtse). We now restrict the analysis to workers with info on years since last studying.

```{r fit31, eval=T, include=T, echo=F, comment=NA, cache=F}
#x2: workers with info on years since last studying
x2=x[x$ehtse>0 & (as.integer(x$esdtl)==11 | as.integer(x$esdtl)==12 | as.integer(x$esdtl)==17),]
model31=(wages_perc ~ s(ehtse_au) + s(ehtse_anglo) + s(ehtse_others) + sex + ancob3 + edhigh1 + jbmo62 + edcoq3 + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct + other_household)
fit31 = gam(model31, data=x2)
anova(fit31)
plot(fit31, rug=F ,pages=1)
```

Taking away ancob3 from the regression:

```{r fit32, eval=T, include=T, echo=F, comment=NA}
model32=(wages_perc ~ s(ehtse_au) + s(ehtse_anglo) + s(ehtse_others) + sex + edhigh1 + jbmo62 + edcoq3 + edqenr + fmfo61 + fmmo61 + esdtl + mrcurr + hhstate + jbmwpsz + jbocct + other_household)
fit32 = gam(model32, data=x2)
anova(fit32)
plot(fit32, rug=F ,pages=1)
```
