# AU data set


```{r eval=T, include=T, echo=F, comment=NA}
library(readstata13)
setwd("~/Dropbox/Attachments/-.Migrations/")
x=read.dta13("IZA_WB_2014.dta")

x_bkup=x
#cat_var = c()
#x[,cat_var] = as.factor(x[,cat_var])

no_sign_digits = 1 #in occupation 4 digits code
x$cur_occ = as.factor(substr(as.character(x$cur_occ),1,no_sign_digits))
x$hfqu = as.factor(x$hfqu)
#attr(x,"label.table")
```

Covariates:
  * sex (1: Male; 2: Female)
* age
* visacat
* marstat
* $c_hfqu0
1: Australia & New Zealand                  
2: UK, Ireland, US & Canada                                    
3: EU/EEA 
4: South, East and South East Asia & Oceania                                     
5: Other 
* $intst2
1: New South Wales & ACT                     
2: Victoria                   
3: Queensland   
4: South Australia & Tasmania 
5: Northern & Western Australia                     
6: Tasmania           
7: Northern Territory                       
8: A.C.T. 
* $hfqu
1: Higher degree               
2: Post graduate diploma       
3: Bachelor degree or equivalent 
4: Tech/prof qual  diploma/certificate                               
5: Trade       
6: 12 or more years of schooling 
7: 10-11 years of schooling              
8: 7-9 years of schooling       
9: 6 or fewer years of schooling 
88: Other 
* $hfqu3
1: Higher degree               
2: Post graduate diploma       
3: Bachelor degree or equivalent 
4: Tech/prof qual  diploma/certificate                               
5: Trade       
8: Not stated/Refused 
9: None of these 
* $qu_cur
1: Higher degree               
2: Post graduate diploma       
3: Bachelor degree or equivalent 
4: Tech/prof qual  diploma/certificate                               
5: Trade       
6: Other certificate 
7: Other
8: Not stated/Refused 
9: Dont know
* $occ_cr, $oc_hc: occupation, generic
* $as_hfqu, $occfhc, $cur_occ, $as_qucur: occupation, very detailed --->> use $cur_occ
* $cur_work: "do you currently work?"
0: Language Difficulties                   
1: Yes                    
2: No    
8: Not stated/Refused                    
9: dk 
* $cit_stat
1: Australian Citizen            
2: Applied        
3: Not Applied 
8: Not stated/Refused                 
9 :dk 
* $marstat
1: Married          
2: Separated           
3: Divorced            
4: Widowed      
5: Never Married 
8: Not stated/Refused                 
9: dk 

Dependent variable:
  * wincwage
0: None, Nil     
1: $1 to 57 a week    
2: $58 to 96 a week   
3: $97 to 154 a week  
4: $155 to 230 a week  
5: $231 to 308 a week 
6: $309 to 385 a week 
7: $386 to 481 a week  
8: $482 to 577 a week  
9: $578 to 673 a week  
10: $674 to 769 a week  
11: $770 to 961 a week 
12: $962 or more a week  
98: Not stated/Refused                  
99: DK 
$pcHHinc
0: None, Nill       
1: 1 to 8,000 AUD a year  
2: 8,001 to 16,000 AUD a year 
3: 16,001 to 25,000 AUD a year 
4: 25,001 to 35,000 AUD a year 
5: 35,001 to 50,000 AUD a year   
6: 50,001 AUD or more a year          
8: Not stated/Refused 
$HHinc
0: None, Nill       
1: 1 to 8,000 AUD a year  
2: 8,001 to 16,000 AUD a year 
3: 16,001 to 25,000 AUD a year 
4: 25,001 to 35,000 AUD a year 
5: 35,001 to 50,000 AUD a year   
6: 50,001 AUD or more a year          
8: Not stated/Refused 

The model:
  Each individual of country j has a wage w_i given a few covariates (sex, age, etc.) and education level (`r edu=x$hfqu`) and job (`r occ=x$cur_occ`). edu and occ must be coupled in factors, but there is the problem of massive missing observations: 18449 have missing values for edu or occ (and 7808 for both of them).
$cur_act is the occupation variable to use (putting together workers/unemployed and non-labour-force). $cur_occ is to be used in the workers category of cur_act.

Keep:
  form_id
intst
visa_mj
sex
age
marsstat
noHHm
nch_res
nrelAU
cobirth
ac_12fhc
imm_y
visAU1
hfqu: reported only once for each form_id,
[see:
   ```{r eval=T, include=T, echo=F, comment=NA}
 ids = unique(x$form_id)
 n_ids = length(ids)
 notNA = NULL
 for (i in 1:n_ids){
   rows=which(x$form_id==ids[i])
   notNA=c(notNA,sum(!is.na(x$hfqu[rows])))
 }
 ```
 ]
but not always in the first interview. Right? <-- Ask Max
If yes, change in this way:
  ```{r eval=T, include=T, echo=F, comment=NA}
ids = unique(x$form_id)
n_ids = length(ids)
for (i in 1:n_ids){
  rows_NA = which(x$form_id==ids[i] & is.na(x$hfqu))
  rows_notNA = which(x$form_id==ids[i] & !is.na(x$hfqu))
  x$hfqu[rows_NA]=x$hfqu[rows_notNA]
}
```
cur_act: can change for same form_id. A pattern like 5, 7, NA means... And NA, 8, NA? <-- Ask Max
cur_occ
valfun1 or logvalfun
car1 or car
ileng
qual_as
occ_hc
wincwage
oved_cr
hours
al06
WAVE: 5 waves (1-3: cohort 1; 4-5: cohort 2) [see table(x$WAVE,x$cohort)]
COHORT
stayjob
t1d
ab22

Possibly useful commands:
  
  > aa=attr(x,"label.table")
> ii=which(names(aa)=='wincwage')
> aa[ii]
> xinds=which(x$cur_occ>=1301 & x$cur_occ<=1399)
> attr(x[xinds,],"label.table")
> xsub=x[x$WAVE==1 | x$WAVE==4,]
> qu=xsub$hfqu
> occ=xsub$cur_occ
> length(which(is.na(qu) & is.na(occ)))

```{r eval=T, include=T, echo=F, comment=NA}
model1=(wincwage ~ WAVE + COHORT + sex + age + hfqu*cur_occ + (1|form_id))
fit1 = lm(model, data=x)
summary(fit)

library(mgcv)
fit2 = gam(wincwage ~ WAVE + COHORT + sex + age + hfqu*cur_occ, data=x)
fit2s = gam(wincwage ~ COHORT + sex + age + hfqu*cur_occ + s(WAVE, k=3), data=x)

fit2r = gamm(wincwage ~ WAVE + COHORT + sex + age + hfqu*cur_occ, random=list(form_id=~1), data=x)
```


# IZA2

```{r eval=T, include=T, echo=F, comment=NA}
library(readstata13)
setwd("~/Dropbox/dropbox mq/Maurizio/")
x=read.dta13("MAURIZIO_IZA_WB_2014.dta")
no_sign_digits = 1 #in occupation 4 digits code
x$cur_occ = as.factor(substr(as.character(x$cur_occ),1,no_sign_digits))
x$hfqu = as.factor(x$hfqu)
```

